Adopting agile in the large: motivations, success factors and challenges

### 2014-05-14 journal/article.tex
Editing: Challenges

@@

@@

To start a transformation the organization needs an initial push. However,
people are usually disincluned to change, and it is likely that resistance
will emerge in any change. 
% Change resistance is a central challenge in any organizational transformation.
People will not be willing to change unless there are good reasons, and the
change is preceived easy enough. It is difficult to attain a buy-in for a
change, and even organizations that have a flexible culture will face resistance
[P10]. It should be expected that not everyone will be willing to change, and
some employees will never adapt to the change [P6]. In one case it was reported
that objections from team members was a major reason for loss in time and
productivity [P29].

On a team level numoerous reasons for change resistance were given. For
instance, Fecarotta described the organization of Boeing as risk-averse and
cautious which slows down any change [P14]. The organization was also described
having a deeply rooted status-quo because of high employee retention, which
further hindered change. In some cases it was reported that people worried about
new roles and responsibilities that agile would possibly bring [P44]. For
instance, testers were worried that they would need to take on cross functional
tasks, which would be outside their area of expertise [P18]. Yet another reason
for change resistance was the move from own offices to team areas [P22]. People
were also feeling being monitored because of the increased level of interaction
within the team and between project stakeholders [P45]. One study reported the
cultural change required for middle management as the biggest cause of
resistance [P2].

In these cases change resistance was mitigated for instance by education [P45],
and one-to-one discussions [P22]. Change resistance was also softened by using
first only teams willing to try agile, so that insights could be gathered to
ease the continued transformation [P45].

@@

The problem of tension between management and development may also be opposite.
In cases where the change was emerging bottom up the reluctancy to change was
not in the team members but management. This caused that significant
organizational change could not be raised above the team level.
For instance, Atlas [P3] describes how it would have required executive support
to integrate the product management office and separate quality assurance groups
in the agile process. In this case, when managers owning projects with many
teams were not involved in the transformation, the agile model was restrained
from spreading beyond team level. Lack of middle management support was also
seen as the most serious shortcoming by Spayd [P49]. Middle management is in a
position to undermine the entire transformation, and may do so if they do not
participate in, and thus understand, the agile method. Agile brings changes to
some management roles [P2], and lack of understanding of agile development will
leave managers feeling left out [P6]. Still one problem is that if management
does not speak out a goal towards using agile, developers feel that the agile
methods may be replaced by something else at any time [P12].

\subsubsection{Other organizational issues}

@@

\subsubsection{Falling back to old model}

@@

A challenge most organizations came across was that an agile model was difficult
implement. An experienced software team may do well in training, but when the
time comes to apply agile techniques in practice the team may get lost [P21].
Agile was also seen as a solution to all problems [P9], success was declared
prematurely [P49], and expectations created by successful early experiments were
not fullfilled [P38]. Disapointment was granted when these hopes were not
fullfilled and implementing an agile model was found to be hard and time
consuming. The difficulty in implementing agile is evident in misconceptions of
the agile model, when by-the-book implementations prove to be insufficient, and
when the agile methods were applied badly to the organization.

@@

Even if there was no misunderstandings the correct agile model was found hard to
implement. Several cases describe how an agile model was hard to learn from
literature [P21, P27], especially when it comes to implementing it on a large
scale [P13]. As Beavers [P5] writes: ``There simply was not a manual or document
where we could find easy answers on how to do things''.
Schnitter and Mackert [P45] report that theoretical considerations on how to
scale up the agile methos were not good enough, and that product managers and
architects struggled when several Scrum teams were working concurrently.
Another case concluded that all practices prescribd in XP did not fit enterprise
scale development, and some practices required customization [P49]. It would not
be good to prescribe practices with a check list as it would impose a single
model in an organization with several individual teams and projects [P41].

Further indications were given that the advice given in literature will not
ferry all the way. Benefield [P6] describes how it was dificult finding a
balance between prescribing a by-the-book implementation which may put people
off, and giving too much freedom in the agile methods which may weaken core
practices.
The reality where the practices needed to be applied was described as
messy in comparison to the ideal circumstances presented in literature, which
made it difficult to choose an initial model and gain buy-in for it [P10].
Difficulties in choosing the model to start with were also due to variances
perceived in the agile approaches suggested in literature [P30].

@@

The difficulty of agile and misunderstandings were also evident in cases where
the methods were customized badly. As by-the-book implementation often was not
feasible the agile method was tailored to suit the needs of the organization.
However, in some cases tailoring meant skipping some practices, which lead into
problems. In one case certain individuals were allowed to ignore some core
elements of Scrum, which came to turn the teams decision making into a variant
of command and control [P9].
Lewis and Neher [P29] describe how there was temptation to strip some practices
and enhance others. Previous attempts had proven that one of the reasons early
attempts of agile had failed was deviations from the process, because of which
the agile mindset did not take root [P29]. In a bad implementation customization
may lead to the team taking only practices that reflect their current needs, and
therefore failing to achieve a real change in process and mindset [P9, P46].
Another example of bad customization that counteracts change is replacing agile
vocabulary in favor of familiar terminology from the old model, because
introducing new language will also create new ways of thinking [P44].

@@

% Coordinating between many teams

%% ?? -->
% Include this ???
% Large organization put agile methods under the test of scaling up, and problems
% were encountered. One case reported use of agile across hundreds of teams, but
% management could not get a roll up of the progress without having strict rules
% on how status was reported [P48]. Schnitter and Mackert [P45] report that a
% maximum workable size in a project managed by Scrum was 130 people including 7
% development teams.

In the agile model it was hard for managers to get an overview of the progress
of the release [P17]. The status of features was hard to tell because different
aspects of features were sepatated across teams [P17]. Smits and Rilliet [P48]
report that providing executive overview of the development programs was limited
by the tools used to manage agile projects.

Some organizations attempted to initially create a model where teams operate
autonomously, which is well in line with agile principles. However, a number of
issues arose from independence. For instance, teams had to balance between the
team's own goals and broader goals of the entire organization, and some teams
chose focus on their own goal [P7]. Coordination was obstructed by independent
teams that did not respect the larger context they were working in [P13].
Coordination was also difficult when different teams were working for different
customers but the applications being built were interdependent [P29]. One case
reported that allowing teams to have different sprint durations created delays
in delivery [P51].

Further problems were encountered when scaling up agile over many sites.
Operating in a distributed environment makes communication more difficult, which
causes problems with agile when frequent and effective communication needs to be
achieved. Distribution had negative effects, such as missing kick-off meetings,
reduced feeling of proximity when telecommunication is necessary, and difficulty
in arranging frequent meetings due to time zone differences [P45].
Agile project management was also problematic in a distributed setting. In a
waterfall model separate parts of projects could be isolated to different sites,
but the agile model does not allow such strict slicing of projects [P29].
In one case it was simply admitted that a distributed organization will impose
additional burden on communication and require additional care in the process,
but distribution and agile would still be used together [P48].

Moore and Spens [P34] describe the progression of the above mentioned
coordination problems. At first cross-team dependencies were attempted to be
reduced, but it became evident that the dependencies were an inherent part of
the project. A traditional approach to managing dependencies created excess
work, reduced independence and flexibility of teams, and created contract based
and adversarial relationships within the organization. There was some success
when up to 5 teams were coordinated using Scrum-of-Scrums, but that model could
not be scaled to a global level. The main problem was teams that were thinking
too much inside the team walls. When the new practices were scaled up the
communication channels narrowed and created communication breakdowns. Problems
with collaboration, dependencies, and integration were solved with presnce of
individuals with an proactive and open mindset towards working over team
boundaries. It was concluded that a balance is needed between completing new
stories from the team backlog and maintaining overall stability of the
application. [P34]

Many studies presented solutions for the inter-team challenges, which also
highlight the challenges themselves. The general model of solving inter-team
problems was introducing an additional layer of coordination, such as a
Scrum-of-Scrums team [P43], a similar product team [P45], or an entire
customized release framework [P13]. In one case integration problems were
controlled by prioritizing code quality and maintaining existing functionality
over new features [P13]. Finally, the choice of tools was reported to have an
impact on the coordination effort [P17].

@@

In most cases a large organizational transformation proceeded gradually, and
during the process it was possible that the new agile model was used in parallel
with the old methods. The arrangement of using the old and new models side by
side was seen as problematic [P1, P26], and causing tensions on all organizational
levels [P10]. Ongoing projects had been set up with previous development
methods, and the agile methods had to be arranged to fit with them [P8, P29].
A particular problem in sychronizing work between projects using different
models was that in Scrum the software design was fleshed out over time as
sprints progressed, but the waterfall model would require detailed design to be
locked early in the project [P29].

@@

@@

% --

% This where?

Some cases reported that development efforts were controlled on a top level by a
project management office (PMO). The PMO was described as entrenched in a
waterfall paradigm, and the top-level rigidity was causing friction in the
agile adoption [P3, P6]. The PMO was seen as a hub and bottleneck controlling
all aspects of projects, and its central structure had to be dismantled for
the organization to become agile [P46].

>>>>

The challenges presented in this section are factors that primary studies
described hindering the progress of the transformation.
Some of the challenges presented here were not explicitly presented as
transformation challenges, but rather as general difficulties in agile
imlementation. These issues are still included as creating an agile
implementation can be seen as a part of the transformation process.
Some solutions for challenges are also presented, but they should be considered
as ways of describing the challenges rather than direct success factors.

@@

% Change resistance
% Lack of investment in transformation
% Agile is found to be hard
% Large organization issues
% Coordinating work between many teams
% Integrating other functions than development to agile
% Other challenges in transformation (zealots, falling back)

@@

To start a transformation the organization needs an initial push. However,
people are usually disinclined to change, and it any organizational change
should anticipate some resistance. People will not be willing to change unless
there are good reasons, and the change is preceived easy enough. It is difficult
to attain a buy-in for a change, and even organizations that have a flexible
culture will face resistance [P10]. It should be expected that not everyone will
be willing to change, and some employees will never adapt to the change [P6].
During a transformation period objections to change may become a major reason
for loss in time and productivity [P29].

On a team level numoerous reasons for change resistance were given. For
instance, Fecarotta [P14] described the organization of Boeing as risk-averse
and cautious which slows down any change. The organization was also described
having a deeply rooted status-quo because of high employee retention, which
further hindered change [P14]. In some cases it was reported that people worried
about new roles and responsibilities that agile would possibly bring [P44]. For
instance, testers were worried that they would need to take on cross functional
tasks, which would be outside their area of expertise [P18]. Yet another reason
for change resistance was the move from own offices to team areas [P22]. People
were also feeling being monitored because of the increased level of interaction
within the team and between project stakeholders [P45]. One study reported the
cultural change required for middle management as the biggest cause of
resistance [P2].

In these cases change resistance was mitigated for instance by education [P45]
and one-to-one discussions [P22]. Change resistance was also softened by first
engaging only teams willing to try agile, so that insights could be gathered to
ease the continued transformation [P45].

@@

The direction in which a tension between management and development is built may
also be opposite. In cases where the change was emerging bottom up the
reluctancy to change was not in the team members but management, causing that
significant organizational change could not be raised above the team level. For
instance, Atlas [P3] describes how it would have required executive support to
integrate the product management office and separate quality assurance groups in
the agile process. In this case, when managers owning projects with many teams
were not involved in the transformation, the agile model was restrained from
spreading beyond team level [P3]. Lack of middle management support was also
seen as the most serious shortcoming by Spayd [P49]. Middle management is in a
position to undermine the entire transformation, and may do so if they do not
participate in, and thus understand, the agile method. Agile brings changes to
some management roles [P2], and lack of understanding of agile development will
leave managers feeling left out [P6]. Still one problem is that if management
does not speak out a goal towards using agile, developers feel that the agile
methods may be replaced by something else at any time [P12].

\subsubsection{Lack of investment in transformation}

It is clear that changing the way of working will create some costs for the
organization. If there is no intention to allow resources for the transformation
problems will ensue. Typical examples on this was lack of investment in
training, and that people were kept allocated to tasks from the original
arrangement. In some cases the transformation was impeded when work spaces could
not be updated.

@@

@@

A challenge most organizations came across was that an agile model was difficult
implement. An experienced software team may do well in training, but when the
time comes to apply agile techniques in practice the team may get lost [P21].
Agile was also thought being a solution to all problems [P9], success was
declared prematurely [P49], and expectations created by successful early
experiments were not fullfilled [P38]. Disapointment was granted when these
hopes were not fullfilled, but instead implementing the agile model was found
to be hard and time consuming. The difficulty in implementing agile is evident
in misconceptions of the agile model, when by-the-book implementations prove to
be insufficient, and when the agile methods were applied badly to the
organization.

@@

Even if there was no misunderstandings the correct agile model was found hard to
implement. Several cases describe how an agile model was hard to learn from
literature [P21, P27], especially when it comes to implementing it on a large
scale [P13]. As Beavers [P5] writes: ``There simply was not a manual or document
where we could find easy answers on how to do things''.
Schnitter and Mackert [P45] report that theoretical considerations on how to
scale up the agile methos were not good enough, and that product managers and
architects struggled when several Scrum teams were working concurrently.
Another case concluded that all practices prescribd in XP did not fit enterprise
scale development, and some practices required customization [P49]. The model of
agile should not be prescribed as a check list, as it would impose a single
model in an organization with several individual teams and projects [P41].

% Check this paragraph
Further indications were given that the advice given in literature will not
ferry all the way. Benefield [P6] describes how it was dificult finding a
balance between prescribing a by-the-book implementation which may put people
off, and giving too much freedom in the agile methods which may weaken core
practices.
The reality where the practices needed to be applied was described as
messy in comparison to the ideal circumstances presented in literature, which
made it difficult to choose an initial model and gain buy-in for it [P10].
Difficulties in choosing the model to start with were also due to variances
perceived in the agile approaches suggested in literature [P30].

@@

The difficulty of agile and misunderstandings were also evident in cases where
the methods were customized badly. As by-the-book implementation often was not
feasible the agile method was tailored to suit the needs of the organization.
However, in some cases tailoring meant skipping some practices, which lead into
problems. In one case certain individuals were allowed to ignore some core
elements of Scrum, which came to turn the teams decision making into a variant
of command and control [P9].
Lewis and Neher [P29] describe how there was temptation to strip some practices
and enhance others. Previous attempts had proven that one of the reasons for
agile implementations to fail was deviations from the process, because of which
the agile mindset did not take root [P29]. A bad customization may lead to the
team taking only practices that reflect their current needs, and therefore
failing to achieve a real change in process and mindset [P9, P46].
Another example of bad customization that counteracts change is replacing agile
vocabulary in favor of familiar terminology from the old model [P44]. A benefit
of introducing new vocabulary is that it may stimulate new ways of thinking
[P44].

\subsubsection{Coordinating work between many teams}

In large software projects work must inevitably be distributed between many
teams. One of the most prominent transformation challenges was the difficulty in
creating a model to coordinate work between many teams.
% --> Present the different aspects of this topic?

% Independence of teams

Some organizations attempted to initially create a model where teams operate
autonomously, which is well in line with agile principles. However, a number of
issues arose from independence. For instance, teams needed to find a balance
between the their own goals and broader goals of the entire organization, but
some teams chose to focus only on their own goal [P7]. Independent teams that
did not respect the larger context they were working in obstructed coordination
[P13]. Coordination was also difficult when teams were working independently for
different customers but the applications being built were interdependent [P29].
One case reported that allowing teams to have different sprint durations created
delays in delivery [P51].

Further coordination problems were encountered when scaling up agile over many
sites. Operating in a distributed environment makes communication more
difficult, and causes problems when frequent and effective communication needs
to be achieved. Distribution had negative effects, such as missing kick-off
meetings, reduced feeling of proximity when telecommunication is necessary, and
difficulty in arranging frequent meetings due to time zone differences [P45].
Agile project management was also problematic in a distributed setting. In a
waterfall model separate parts of projects could be isolated to different sites,
but the agile model does not allow such strict slicing of projects [P29]. In one
case it was simply admitted that a distributed organization will impose
additional burden on communication and require additional care in the process,
but distribution and agile would still be used together [P48].

Moore and Spens [P34] describe the progression of the above mentioned
coordination problems. At first cross-team dependencies were attempted to be
reduced, but it became evident that the dependencies were an inherent part of
the project. A traditional approach to managing dependencies created excess
work, reduced independence and flexibility of teams, and created contract based
and adversarial relationships within the organization. There was some success
when up to 5 teams were coordinated using Scrum-of-Scrums, but that model could
not be scaled to a global level. The main problem was teams that were thinking
too much inside the team walls. When the new practices were scaled up the
communication channels narrowed and created communication breakdowns. Problems
with collaboration, dependencies, and integration were solved in the presnce of
individuals having a proactive and open mindset towards working over team
boundaries. It was concluded that a balance is needed between completing new
stories from the team backlog and maintaining overall stability of the
application. [P34]

Many studies presented solutions for the inter-team challenges, which also
highlight the nature of the challenges themselves. The general model of solving
inter-team problems was introducing an additional layer of coordination, such as
a Scrum-of-Scrums team [P43], a similar product team [P45], or an entire
customized release framework [P13]. In one case integration problems were
controlled by prioritizing code quality and maintaining existing functionality
over new features [P13]. Finally, the choice of tools was reported to have an
impact on the coordination effort [P17].

@@

%% ?? -->
% Include this ???
% Large organization put agile methods under the test of scaling up, and problems
% were encountered.
% In the agile model it was hard for managers to get an overview of the progress
% of the release [P17]. The status of features was hard to tell because different
% aspects of features were sepatated across teams [P17].
% One case reported use of agile across hundreds of teams, but management could
% not get a roll up of the progress without having strict rules on how status was
% reported [P48]. Schnitter and Mackert [P45] report that a maximum workable size
% in a project managed by Scrum was 130 people including 7 development teams.
% % Smits and Rilliet [P48] report that providing executive overview of the
% % development programs was limited by the tools used to manage agile projects.

@@

In most cases a large organizational transformation proceeded gradually, and
during the process it was possible that the new agile model was used in parallel
with the old methods. The arrangement of using the old and new models side by
side was seen as problematic [P1, P26], and causing tensions on all organizational
levels [P10]. Ongoing projects had been set up with previous development
methods, and the agile methods needed to be arranged to fit with them [P8, P29].
A particular problem in sychronizing work between projects using different
models was that in Scrum the software design was fleshed out over time as
sprints progressed, but the waterfall model would require detailed design to be
locked early in the project [P29].

@@

% This where?
Some cases reported that development efforts were controlled on a top level by a
project management office (PMO). The PMO was described as entrenched in a
waterfall paradigm, and the top-level rigidity was causing friction in the
agile adoption [P3, P6]. The PMO was seen as a hub and bottleneck controlling
all aspects of projects, and its central structure had to be dismantled for
the organization to become agile [P46].

@@

\subsubsection{Other challenges in transformation}

% Falling back to old model
